---
title: "R Notebook"
output: html_notebook
---

```{r}
library(pracma)
library(extraDistr)
```





$t$-th power moment $K_t$:

$$K_t(D)=\sum_{1\le i < j\le N}[\delta_{ij}(D)]^t/(N(N-1)/2)$$
其中 $\delta_{ij}(D)$ 是 $D$ 中第 $i$ 和第 $j$ 列的相同元素的個數。

```{r}
# random 12x16 SSD

design <- matrix(rbinom(12*16, 1, 0.5), 12, 16)*2-1

K_function <- function(design, t=2) {
  N <- nrow(design)
  delta <- matrix(0, N, N)
  for (i in 1:(N-1)) {
    for (j in (i+1):N) {
      delta[i, j] <- sum(design[i, ] == design[j, ])
    }
  }
  return(sum(delta[delta != 0]^t) / (N*(N-1)/2))
}

K_function(design, 2)
```

$$A_t(D)=E(J_t^2)/N^2$$

```{r}
A_function <- function(design, t=2) {
  n <- ncol(design)
  N <- nrow(design)
  if (t == 1) {
    return(sum(apply(design, 2, sum)^2) / N^2)
  } else {
    sum_products <- 0
    comb <- combn(n, t)
    for (i in 1:choose(n, t)) {
      sum_products <- sum_products + sum(apply(design[, comb[, i]], 1, prod))^2
    }
    return(sum_products / N^2)
  }
}

A_function(design, 2)
```

在所有因子 level 數都為 2 的情況下:

\begin{aligned}
  K_2&=\frac{[2A_2+2nA_1+n(n+1)]N-4n^2}{4(N-1)}\\
  K_3&=\frac{[6A_3+6nA_2+(3n^2+3n-2)A_1+n^3+3n^2]N-8n^3}{8(N-1)}
\end{aligned}

$A_1=0$ if the design is balance.

\begin{aligned}
  A_2&=\frac{2(N-1)}{N}K_2-\frac{n(n+1)N-4n^2}{2N}\\
  A_3&=\frac{4(N-1)}{3N}K_3-\frac{2n(N-1)}{N}K_2+\frac{n^3N-2n^3}{3N}
\end{aligned}

```{r}
n <- ncol(design)
N <- nrow(design)

K_function(design, 2)
((2*A_function(design, 2) + 2*n*A_function(design, 1) + n*(n+1))*N - 4*n^2) / (4*(N-1))

K_function(design, 3)
((6*A_function(design, 3) + 6*n*A_function(design, 2) + (3*n^2+3*n-2)*A_function(design, 1) + n^3 + 3*n^2)*N - 8*n^3) / (8*(N-1))


# a random 12x16 balance SSD

design <- cbind(rep(c(-1, 1), each=6)[randperm(12, 12)])
for (i in 1:15) {
  design <- cbind(design, rep(c(-1, 1), each=6)[randperm(12, 12)])
}

A_function(design, 2)
2*(N-1)/N*K_function(design, 2) - (n*(n+1)*N-4*n^2)/(2*N)
A_function(design, 3)
4*(N-1)/(3*N)*K_function(design, 3) - 2*n*(N-1)/N*K_function(design, 2) + 
  (n^3*N-2*n^3)/(3*N)
```

$$K_1=\frac{n(N-2)}{2(N-1)}=\sum_{1\le i < j\le N}\delta_{ij}(D)/{N\choose 2}$$

```{r}
design <- cbind(rep(c(-1, 1), each=6)[randperm(12, 12)])
for (i in 1:14) {
  design <- cbind(design, rep(c(-1, 1), each=6)[randperm(12, 12)])
}

K_function(design, 1)*choose(nrow(design), 2)

N <- nrow(design)
delta <- matrix(0, N, N)
for (i in 1:(N-1)) {
  for (j in (i+1):N) {
    delta[i, j] <- sum(design[i, ] == design[j, ])
  }
}

table(delta[upper.tri(delta)])

```

$N=16, n=32$ SSD:  
to minimize $\lambda A_2+(1-\lambda)A_3$，$\lambda=\frac{10}{11}$

\begin{aligned}
\lambda A_2+(1-\lambda)A_3
&=\frac{2(N-1)}{N}[n+\lambda(1-n)]K_2+(1-\lambda)\frac{4(N-1)}{3N}K_3+c(N, n, \lambda)\\
&\propto\lambda E(J_2^2)\times {n\choose 2}+(1-\lambda)E(J_3^2) {n\choose 3}
\end{aligned}


```{r}
CreateSSD <- function(lambda = 0.5, N=12, n=16) {
  design <- cbind(rep(c(-1, 1), each=N/2), rep(c(-1, 1, -1, 1), each=N/4))
  # design <- FFD2(4)
  for (column in 1:(n-2)) {
    last_added <- Inf
    for (try in 1:100) {
      # randomly generate a balance length N vector
      col_added <- rep(c(-1, 1), each=(N/2))[randperm(N, N)]
      delta <- matrix(0, nrow(design), nrow(design))
      for (i in 1:(nrow(design)-1)) {
        for (j in (i+1):nrow(design)) {
          delta[i, j] <- sum(cbind(design, col_added)[i, ] == cbind(design, col_added)[j, ])
        }
      }
      
      K2 <- sum(delta[upper.tri(delta)]^2)
      K3 <- sum(delta[upper.tri(delta)]^3)
      last_value <- (2*(N-1)*(n+lambda*(1-n))/N)*K2 + (1-lambda)*(4*(N-1)/3/N)*K3
      
      
      
      while (T) {
        pos_place <- which(col_added==1)
        neg_place <- which(col_added==-1)
        value_added <- matrix(0, N/2, N/2)
        i <- 0
        for (place1 in pos_place) {
          i <- i + 1
          j <- 0
          for (place2 in neg_place) {
            j <- j + 1
            
            delta_star <- delta
            delta_star[place1, neg_place[-j]] <- delta_star[place1, neg_place[-j]] + 1
            delta_star[place1, pos_place[-i]] <- delta_star[place1, pos_place[-i]] - 1
            delta_star[place2, pos_place[-i]] <- delta_star[place2, pos_place[-i]] + 1
            delta_star[place2, neg_place[-j]] <- delta_star[place2, neg_place[-j]] - 1
            delta_star <- (delta_star[upper.tri(delta_star)]+t(delta_star)[upper.tri(delta_star)])
            K2 <- sum(delta_star^2)
            K3 <- sum(delta_star^3)
            
            value_added[i, j] <- (2*(N-1)*(n+lambda*(1-n))/N)*K2 + (1-lambda)*(4*(N-1)/3/N)*K3
          }
        }
        
        # if no improvement, then break
        if (min(value_added) >= last_value) {
          break
        } else {
          exchange_places <- which(value_added == min(value_added), arr.ind=TRUE)[1, ]
          col_added[c(pos_place[exchange_places[1]], neg_place[exchange_places[2]])] <-
            col_added[c(neg_place[exchange_places[2]], pos_place[exchange_places[1]])]
        }
        last_value <- min(value_added)
      }
      if (last_value < last_added) {
        last_added <- last_value
        best_col_added <- col_added
      }
    }
    
    # append design matrix
    design <- cbind(design, best_col_added)
  }
  colnames(design) <- 1:ncol(design)
  return(design)
}

lambda <- 10/11
start_time <- Sys.time()
design1 <- CreateSSD(lambda, 16, 32)
Sys.time() - start_time

dim(design1)
A_function(design1, 2)
A_function(design1, 3)

lambda*J_function(design1, 2) + (1-lambda)*J_function(design1, 3)

K_function(design1, 2)
K_function(design1, 3)
```


```{r}
N <- 16
delta <- matrix(0, N, N)
for (i in 1:(N-1)) {
  for (j in (i+1):N) {
    delta[i, j] <- sum(design1[i, ] == design1[j, ])
  }
}
delta
```



```{r}
CreateSSD <- function(lambda = 0.5) {
  design <- cbind(rep(c(-1, 1), each=N/2), rep(c(-1, 1, -1, 1), each=N/4))
  FI2 <- matrix(0, nrow(design), choose(ncol(design), 2))
  count <- 0
  for (i in 1:(ncol(design)-1)){
    for (j in (i+1):ncol(design)){
      count <- count + 1
      FI2[, count] <- design[, i]*design[, j]
    }
  }
  
  J23 <- 0
  for (column in 1:30) {
    last_added <- Inf
    for (i in 1:100) {
      # randomly generate a balance length 16 vector
      col_added <- rep(c(-1, 1), each=8)[randperm(16, 16)]
      last_J23_plus <- lambda*sum((t(design) %*% col_added)^2) + 
        (1-lambda) * sum((t(FI2) %*% col_added)^2)
      while (T) {
        pos_place <- which(col_added==1)
        neg_place <- which(col_added==-1)
        
        value_added <- matrix(0, 8, 8)
        j <- 0
        for (place1 in pos_place) {
          j <- j + 1
          k <- 0
          for (place2 in neg_place) {
            k <- k + 1
            col_added_exchange <- col_added
            col_added_exchange[c(place1, place2)] <- col_added_exchange[c(place2, place1)]
            J2_plus <- sum((t(design) %*% col_added_exchange)^2)
            J3_plus <- sum((t(FI2) %*% col_added_exchange)^2)
            value_added[j, k] <- lambda*J2_plus + (1-lambda)*J3_plus
          }
        }
        
        # if no improvement, then break
        if (min(value_added) >= last_J23_plus) {
          break
        } else {
          exchange_places <- which(value_added == min(value_added), arr.ind=TRUE)[1, ]
          col_added[c(pos_place[exchange_places[1]], neg_place[exchange_places[2]])] <-
            col_added[c(neg_place[exchange_places[2]], pos_place[exchange_places[1]])]
        }
        last_J23_plus <- min(value_added)
      }
      if (last_J23_plus < last_added) {
        last_added <- last_J23_plus
        best_col_added <- col_added
      }
    }
    
    # append FI2
    for (m in 1:ncol(design)) {
      FI2 <- cbind(FI2, best_col_added * design[, m])
    }
    # append design matrix
    design <- cbind(design, best_col_added)
    J23 <- J23 + last_added
  }
  return(design)
}

start_time <- Sys.time()
design2 <- CreateSSD(lambda = lambda)
Sys.time() - start_time

dim(design2)

A_function(design2, 2)
A_function(design2, 3)
lambda*J_function(design2, 2) + (1-lambda)*J_function(design2, 3)

K_function(design2, 2)
K_function(design2, 3)
```

```{r}
N <- 16
delta <- matrix(0, N, N)
for (i in 1:(N-1)) {
  for (j in (i+1):N) {
    delta[i, j] <- sum(design2[i, ] == design2[j, ])
  }
}
delta
```

```{r}
cstar <- function(N, n, lambda) {
  (n^3*N-2*n^3)/3/N - lambda*((n^3*N-2*n^3)/3/N + (n*(n+1)*N-4*n^2)/2/N)
}
```


Try to solve the following problem numerically:

Given $x_1, \dots, x_n$ nonnegative integers and $\sum_{i=1}^nx_i=m$, to minimize

$$\theta\sum_{i=1}^nx_i^2+(1-\theta)\sum_{i=1}^nx_i^3$$

$\theta=\frac{\frac{2(N-1)}{N}[n+\lambda(1-n)]}{\frac{2(N-1)}{N}[n+\lambda(1-n)]+(1-\lambda)\frac{4(N-1)}{3N}}$，$x_i$ 代表某兩列 coincidence 數。

```{r}
weighted_target <- function(x, lambda) {
  return(lambda*sum(x^2) + (1-lambda)*sum(x^3))
}

# genetic algorithm

GA <- function(m, n, theta, max_iter=1000, pop_size=100, mutation_rate=0.1) {
  stop_flag <- FALSE
    pop <- matrix(0, pop_size, n)
    for (i in 1:pop_size) {
      pop[i, ] <- rmultinom(1, m, rep(1, n))
    }
    
    record <- numeric(max_iter)
    for (iter in 1:max_iter) {
      # calculate fitness
      fitness <- rep(0, pop_size)
      for (i in 1:pop_size) {
        fitness[i] <- weighted_target(pop[i, ], theta)
      }
      record[iter] <- min(fitness)
      
      # select parents
      parents <- matrix(0, pop_size, n)
      for (i in 1:pop_size) {
        parents[i, ] <- pop[which.min(fitness[sample(1:pop_size, 2, replace=FALSE)]), ]
        # parents[i, ] <- pop[sample(1:pop_size, 2, replace=FALSE), which.min(fitness)]
      }
      
      # crossover
      children <- matrix(0, pop_size, n)
      for (i in 1:pop_size) {
        portion <- runif(1)
        children[i, ] <- 0.5*parents[sample(1:pop_size, 1), ] + 0.5*parents[sample(1:pop_size, 1), ]
        # half of the non-integers round up, the other half round down
        is.fraction <- (children[i, ] - floor(children[i, ])) > 0
        round_up <- sample(which(is.fraction), sum(is.fraction) %/% 2)
        children[i, round_up] <- ceiling(children[i, round_up])
        round_down <- setdiff(which(is.fraction), round_up)
        children[i, round_down] <- floor(children[i, round_down])
        
        if (sum(children[i, ]) != m) {
          print(children[i, ])
          print(iter)
          stop_flag <- TRUE
          break
        }
        
      }
      if (stop_flag) {
        break
      }
      # mutation
      for (i in 1:pop_size) {
        if (runif(1) < mutation_rate) {
          mutate_index <- sample(1:n, 1)
          children[i, mutate_index] <- children[i, mutate_index] + 1
          mutate_index <- sample(which(children[i, ] > 0), 1)
          children[i, mutate_index] <- children[i, mutate_index] - 1
        }
      }
      
      # select survivors
      for (i in 1:pop_size) {
        if (weighted_target(children[i, ], theta) < weighted_target(pop[i, ], theta)) {
          pop[i, ] <- children[i, ]
        }
      }
      
    }
  plot(record[1:max_iter], type="l")
  return(pop[which.min(apply(pop, 1, weight_J, lambda)), ])
}


lambda <- 10/11
lambda_a <- 2*(N-1)*(n+lambda*(1-n))/N
lambda_b <- (1-lambda)*4*(N-1)/3/N
theta <- lambda_a / (lambda_a + lambda_b)

start_time <- Sys.time()
opt <- GA(K_function(design1, 1) * N * (N-1) / 2, choose(N, 2), theta, max_iter = 15000)
print(Sys.time() - start_time)
table(opt)

weighted_target(delta[upper.tri(delta)], theta)
weighted_target(opt, lambda_star)
weighted_target(c(rep(14, 8), rep(15, 112)), theta)
```

Q: 如何確定 opt 能在 design 中出現?
